Baseline centering problem

### 2018-12-17 paper/tangotd.tex

Meanwhile,
the right Bellman equation becomes
\begin{gather}
\transp{\1_{s_1}}\,\Delta^{-1}\psi=\transp{\1_{s_1}}\,\psi+\transp{\1_{s_1}}\,\Delta^{-1}P\psi
\qquad \Longrightarrow
\\
\green(s_1)\cdot\E_{s_2\sim \rho}
f(s_2)\psi(s_2)
\approx \psi(s_1)+ \green(s_1)\cdot\E_{s_2\sim \rho}\E_{s'_2\sim P(s'_2|s_2)}
f(s_2)\psi(s'_2)
\end{gather}
(indeed, 
$P\psi$ is
represented by $\E_{s_2\sim \rho} f(s_2)(P\psi)(s_2)$, and
$(P\psi)(s_2)=\E_{s'_2\sim P(s'_2|s_2)}\psi(s'_2)$).

@@

Nico suggests the references around successor representation, proto-value
functions, byt Marlos Machado.

>>>>

Meanwhile,
the right Bellman equation becomes
\begin{gather}
\transp{\1_{s_1}}\,\Delta^{-1}\psi=\transp{\1_{s_1}}\,\psi+\transp{\1_{s_1}}\,\Delta^{-1}P\psi
\qquad \Longrightarrow
\\
\green(s_1)\cdot\E_{s_2\sim \rho}
f(s_2)\psi(s_2)
\approx \psi(s_1)+ \green(s_1)\cdot\E_{s_2\sim \rho}\E_{s'_2\sim P(s'_2|s_2)}
f(s_2)\psi(s'_2)
\end{gather}
(indeed, 
$P\psi$ is
represented by $\E_{s_2\sim \rho} f(s_2)(P\psi)(s_2)$, and
$(P\psi)(s_2)=\E_{s'_2\sim P(s'_2|s_2)}\psi(s'_2)$). If we view this as
an equation on $f$ for fixed $\green$, this imposes a kind of spatial
regularity of features $f$.

@@

Nico suggests the references around successor representation, proto-value
functions, by Marlos Machado.

\paragraph{Enumerative combinatorics, the middle Bellman equation, and
the Newton method for $\Delta^{-1}$.} As mentioned above, $(\Id-\gamma
P)^{-1}$ counts the number of paths from state $s_1$ to $s_2$, weighted
by their probability and $\gamma^\mathrm{length}$. When observing a new
transition $s\to s'$, the left Bellman equation on $P$ produces new paths
from $s$, by adding the edge $s\to s'$ to all paths from $s'$. Similarly
the right Bellman equation produces new paths to $s'$.

The Newton method for matrix inversion is a classical tool to compute
$\Delta^{-1}$. It consists in iterating
\begin{equation}
A\gets 2A-A\Delta A
\end{equation}
It converges extremely fast if initialized reasonably close to $A^{-1}$ (namely if the spectral
radius of $\Id-AX$ is less than $1$), but it can diverge if badly
initialized. A suitable initialization is $A=\eps\transp{\Delta}$ for
small enough $\eps$. For stochastic estimates $\tilde \Delta$ of $\Delta$ (as we have
here through transitions), the stochastic update version with learning
rate $\eta$ is
\begin{equation}
A\gets A+\eta A-\eta A\tilde \Delta A
\end{equation}

With $\Delta=\Id-\gamma P$, 
this method corresponds to a ``middle'' Bellman equation
\begin{equation}
\Delta^{-1}\gets \Delta^{-1}+\eta\Delta^{-1}-\eta \Delta^{-2}+\eta \gamma \Delta^{-1}P\Delta^{-1}
\end{equation}
Importantly, when a new transition $s\to s'$ is observed from $P$, this equation
produces new paths by connecting all paths that end at $s$ with all paths
starting at $s'$, thus doubling path length (thanks to the $\Delta^{-1}$
on the right and left of $P$). This doubling corresponds to the
quadratic convergence of the Newton method. (However, this method does not
produce new starting points or endpoints.)

Here we do not have access to $\Delta$, but have access to transitions
$s\to s'$, which are samples of
$\rho\Delta$, namely, $\rho\Delta=\E_{s,s'} 1_s\transp{(\1_s-\gamma
\1_{s'})}$. This corresponds to a stochastic version of the Newton
method, with learning rate $\eta\ll 1$,
\begin{equation}
A\gets A+\eta A-\eta A 1_s\transp{(\1_s-\gamma \1_{s'})} A
\end{equation}
Indeed, in expectation over $s\to s'$ this corresponds to $A\gets A+\eta
A-\eta A \rho\Delta A$, which converges to $A=\Delta^{-1}\rho^{-1}$. This
also corresponds to the rank-one update formyla for matrix inversion,
$(D+\eta u\transp{v})^{-1}=D^{-1}-\eta
D^{-1}u\transp{v}D^{-1}+O(\eta^2)$. Namely,
if we update our estimate of $\rho\Delta$ by
$\rho\Delta\gets (1-\eta)\rho\Delta+\eta 1_s\transp{(\1_s-\gamma
\1_{s'})}$ when a transition $s\to s'$ is observed, then
$(\rho\Delta)^{-1}$ gets updated as above.

With $A=\green\cdot \E_{\rho} f$ as above, everything is fully samplable.
It is not clear to me how to mix the left, right and middle Bellman
equations, but the correspondence with path combinatorics and the Newton
method suggests this
might be an important point for the navigation problem.
